---
title: "Amphibian lens transmission: 3-ecology and transmission"
author: "Katie Thomas"
date: "10 Nov 2021"
output:
  html_document:
    theme: cerulean
    toc: TRUE
    toc_float: TRUE
    code_fold: hide
editor_options: 
  chunk_output_type: inline
---

Here, we examine how ecology may shape species differences in short-wavelength lens transmission. 

```{r setup, include=FALSE}

#libraries
library(knitr)
library(kableExtra)
library(plotly)
library(cowplot)
library(AmphiNom)
library(ggtree)
library(geiger)
library(ape)
library(phytools)
library(caper)
library(plyr)
library(tidyverse)

#set rmarkdown options
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.height = 5, fig.width = 8, attr.output='style="max-height: 350px;"') 

```


# Data

First, we import the tidied data compiled from all studies that includes ASW names and ecological details. 

```{r}

#import species data for lens transmission
lens_data <- data.frame(read.csv("../Data/tidy data/lenses_compiled.csv", header=TRUE, na.strings=c("", "NA"))) 

```

Then we subset data to adults measured in this study and previously published studies. We also create subsets for data from this study vs. data from previous studies. 

```{r}

#adults
lens_adults <- lens_data %>% 
  filter(stage %in% c("adult", "unknown"))

#adults-this study
lens_adults.new <- lens_adults %>% 
  filter(source == "this_study")

#adults-literature
lens_adults.pub <- lens_adults %>% 
  filter(source == "literature")

```

Finally, we import the pruned tree. 

```{r}
#Import pruned tree
lens.tree <- read.tree(file = "../Data/tidy data/pruned_tree.txt")
```


# Visualize data

Here, we examine the distribution of lens transmission summary data across the amphibian phylogeny and ecological traits.

## Plot adult sampling onto phylogeny

```{r, results = "hide"}

# prep data -----

#create dataframe of adults with this paper and lit data in two separate columns

#our data
new_cols <- lens_adults.new %>%
  mutate(t450nm_new = t450nm) %>%
  mutate(t50_new = t50) %>%
  mutate(pUVA_new = pUVA) %>%
  select(ASW_names, t450nm_new, t50_new, pUVA_new, hab, act)

#published data
lit_cols <- lens_adults.pub %>%
  mutate(t450nm_lit = t450nm) %>%
  mutate(t50_lit = t50) %>%
  mutate(pUVA_lit = pUVA) %>%
  select(ASW_names, t450nm_lit, t50_lit, pUVA_lit, hab, act)

#merge
adults_sep <- full_join(new_cols, lit_cols, by = c("ASW_names", "hab", "act"))

#prune tree to adult data -----

#make list of taxa to drop (in tree but not in dataset)
drops <- setdiff(lens.tree$tip.label, adults_sep$ASW_names)

#drop unwanted tips from phylogeny
tree.adults <- drop.tip(phy = lens.tree, tip = drops) 

#set rownames of adult dataset
rownames(adults_sep) <- as.character(adults_sep$ASW_names)

#check that phylogeny tips and data match exactly (if they match will return "OK")
name.check(tree.adults, adults_sep)

#resort dataset to the order of tree tip labels
adults_sep <- adults_sep[tree.adults$tip.label, ] 

```

```{r, fig.height=12, fig.width=5, fig.cap="Phylogeny showing adult lens transmission and ecological traits. Non-italicized values show our t50 (unbolded) and %UVA (bolded) measurements for adult species means, while italicized values show species means calculated from published literature."} 

#plot phylogeny with sampling tip labels ----


#color vecctors
col_hab <- c("other" = "grey95",
             "scansorial" = "#009E73")

col_act <- c("diurnal" = "#FFB122",
             "nondiurnal" = "grey95")

#phylogeny
plot.phylo(tree.adults, 
           type = "phylogram", 
           show.tip.label = TRUE, 
           cex = 0.6, #text size
           no.margin = TRUE, 
           use.edge.length = TRUE, 
           edge.width = 2,
           x.lim = 650,
           label.offset = 2) 

#add tip labels for our data (t50)
tiplabels(adults_sep$t50_new,
          cex = 0.6,
          frame = "none",
          bg = "white",
          offset = 195) #size of labels

#add tip labels for published data (t50) 
tiplabels(adults_sep$t50_lit,
          cex = 0.6,
          font = 3,
          frame = "none",
          bg = "white",
          offset = 215) #size of labels

#add tip labels for our data (%UVA)
tiplabels(round(adults_sep$pUVA_new, digits = 0),
          cex = 0.6,
          font = 2,
          frame = "none",
          bg = "white",
          offset = 240) #size of labels

#add tip labels for published data (%UVA)
tiplabels(round(adults_sep$pUVA_lit,digits = 0),
          cex = 0.6,
          font = 4, 
          frame = "none",
          bg = "white",
          offset = 255) #size of labels

#add tip labels for scansorial/not
tiplabels(col = col_hab[as.factor(adults_sep$hab)], #sets color habitat
          pch = 17,
          cex = 0.8,
          offset = 150) 

#add tip labels for diurnal/not
tiplabels(col = col_act[as.factor(adults_sep$act)], #sets color habitat
          pch = 19,
          cex = 0.8,
          offset = 160) 


#add legend for ecology traits
legend(x = 20, y = 25, legend = c("Scansorial", "Diurnal"), 
       col = c("#009E73", "#FFB122"),
       pch = c(17,19),
       cex = 0.6, 
       box.lty = 0, 
       title = "Ecology", 
       title.adj = 0)

```

```{r, results = "hide"}
#export pdf

#pdf
pdf("../Figures/phylogeny_summary.pdf", width = 7, height = 12)

#phylogeny
plot.phylo(tree.adults, 
           type = "phylogram", 
           show.tip.label = TRUE, 
           cex = 0.6, #text size
           no.margin = TRUE, 
           use.edge.length = TRUE, 
           edge.width = 2,
           x.lim = 650,
           label.offset = 2) 

#add tip labels for our data (t50)
tiplabels(adults_sep$t50_new,
          cex = 0.6,
          frame = "none",
          bg = "white",
          offset = 195) #size of labels

#add tip labels for published data (t50) 
tiplabels(adults_sep$t50_lit,
          cex = 0.6,
          font = 3,
          frame = "none",
          bg = "white",
          offset = 215) #size of labels

#add tip labels for our data (%UVA)
tiplabels(round(adults_sep$pUVA_new, digits = 0),
          cex = 0.6,
          font = 2,
          frame = "none",
          bg = "white",
          offset = 240) #size of labels

#add tip labels for published data (%UVA)
tiplabels(round(adults_sep$pUVA_lit,digits = 0),
          cex = 0.6,
          font = 4, 
          frame = "none",
          bg = "white",
          offset = 255) #size of labels

#add tip labels for scansorial/not
tiplabels(col = col_hab[as.factor(adults_sep$hab)], #sets color habitat
          pch = 17,
          cex = 0.8,
          offset = 150) 

#add tip labels for diurnal/not
tiplabels(col = col_act[as.factor(adults_sep$act)], #sets color habitat
          pch = 19,
          cex = 0.8,
          offset = 160) 

#add legend for ecology traits
legend(x = 20, y = 25, legend = c("Scansorial", "Diurnal"), 
       col = c("#009E73", "#FFB122"),
       pch = c(17,19),
       cex = 0.6, 
       box.lty = 0, 
       title = "Ecology", 
       title.adj = 0)

dev.off()

```



## Plots of t50 and %UVA values across ecological states

Here we plot adult lens transmission metrics (t50 and %UVA) across ecological categories to visualize differences (without accounting for phylogeny)

```{r cols, results = "hide"}

# create vectors of colors for habitat
col_hab <- c("scansorial" = "#009E73",
              "other"  = "#636363")

#create vector of colors for activity period
col_act <- c("nondiurnal" = "#636363",
             "diurnal" = "#FFA010")

#create shape vector for source of data
sh_source <- c("literature" = 1, 
               "this_study" = 19)
```


```{r boxplots-eco}

# Scansoriality ------

# boxplot of T50 across habs
boxplot_habs.t50 <- ggplot(data = lens_adults, 
                       aes(x = hab, y = t50)) + 
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.9) + #controls boxes  
 geom_jitter(aes(color = hab, shape = source), size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  scale_color_manual(values = col_hab,
                     name = "Habitat",
                     breaks = c("scansorial", "other"),
                     labels = c("scansorial", "non-scans.")) +
  stat_summary(fun.y = mean, colour = "black", geom = "point", shape = 18, size = 3, show_guide = FALSE) + #controls what stats shown
  scale_shape_manual(values = sh_source,
                     name = "Data",
                     breaks = c("this_study", "literature"),
                     labels = c("this study", "literature"),
                     guide = guide_legend(override.aes = list(color = "black"), fill = "black")) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background +
  scale_x_discrete(labels=c("scansorial" = "scansorial",
                            "other" = "non-scans.")) +
  scale_y_continuous(breaks = c(300,325,350,375,400,425), limits=c(300, 425)) +
  xlab("") +
  ylab("t50 (nm)")

# boxplot of %UVA across habs
boxplot_habs.uva <- ggplot(data = lens_adults, 
                       aes(x = hab, y = pUVA)) + 
  geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.9) + #controls boxes  
  
 geom_jitter(aes(color = hab, shape = source), size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  stat_summary(fun.y = mean, colour = "black", geom = "point", shape = 18, size = 3, show_guide = FALSE) + #controls what stats shown
  scale_color_manual(values = col_hab,
                     name = "Habitat",
                     breaks = c("scansorial", "other"),
                     labels = c("scansorial", "non-scans.")) +
  scale_shape_manual(values = sh_source,
                     name = "Data",
                     breaks = c("this_study", "literature"),
                     labels = c("this study", "literature"),
                     guide = guide_legend(override.aes = list(color = "black"), fill = "black")) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background
  xlab("Scansoriality") +
  scale_x_discrete(labels=c("scansorial" = "scansorial",
                            "other" = "non-scans.")) +
  ylab("UVA (%)") +
  labs(fill = "Data source")

#print plots
plot_grid(boxplot_habs.t50, boxplot_habs.uva,  #list of plots to arrange in grid
           align = 'vh', #align horizontally and vertically
           ncol = 2,
           rel_widths = c(1, 1)) #number of columns in grids

# Diurnality ------

# boxplot of T50 across acts
boxplot_act.t50 <- lens_adults %>%
  mutate(act = fct_relevel(act, "diurnal", "nondiurnal")) %>%
  ggplot(aes(x = act, y = t50)) + 
   geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.9) + #controls boxes  
  #controls what stats shown
 geom_jitter(aes(color = act, shape = source), size = 2, alpha = 0.7, position = position_jitter(0.15)) +
   stat_summary(fun.y = mean, colour = "black", geom = "point", shape = 18, size = 3, show_guide = FALSE) +
  scale_color_manual(values = col_act,
                     name = "Activity",
                     breaks = c("diurnal", "nondiurnal"),
                     labels = c("diurnal", "non-diurnal")) +
  scale_shape_manual(values = sh_source,
                     name = "Data",
                     breaks = c("this_study", "literature"),
                     labels = c("this study", "literature"),
                     guide = guide_legend(override.aes = list(color = "black"), fill = "black")) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background
  scale_x_discrete(labels=c("diurnal" = "diurnal",
                            "nondiurnal" = "non-diurnal")) +
  scale_y_continuous(breaks = c(300,325,350,375,400,425), limits=c(300, 425)) +
  xlab("Diurnality") +
  ylab("t50 (nm)") +
  labs(fill = "Data source")

# boxplot of %UVA across habs
boxplot_act.uva <- lens_adults %>%
  mutate(act = fct_relevel(act, "diurnal", "nondiurnal")) %>%
  ggplot(aes(x = act, y = pUVA)) + 
   geom_boxplot(notch = FALSE, outlier.shape = NA, alpha = 0.9) + #controls boxes  
 geom_jitter(aes(color = act, shape = source), size = 2, alpha = 0.7, position = position_jitter(0.15)) +
  stat_summary(fun.y = mean, colour = "black", geom = "point", shape = 18, size = 3, show_guide = FALSE) +
  scale_color_manual(values = col_act,
                     name = "Activity",
                     breaks = c("diurnal", "nondiurnal"),
                     labels = c("diurnal", "non-diurnal")) +
  scale_shape_manual(values = sh_source,
                     name = "Data",
                     breaks = c("this_study", "literature"),
                     labels = c("this study", "literature"),
                     guide = guide_legend(override.aes = list(color = "black"), fill = "black")) +
  theme(text = element_text(size=14), panel.background = element_blank(), axis.line = element_line(colour = "black"), legend.key = element_rect(fill = NA)) + #controls background
  scale_x_discrete(labels=c("diurnal" = "diurnal",
                            "nondiurnal" = "non-diurnal")) +
  xlab("Diurnality") +
  ylab("UVA (%)") +
  labs(fill = "Data source")

#print plots
plot_grid(boxplot_act.t50, boxplot_act.uva,  #list of plots to arrange in grid
           align = 'vh', #align horizontally and vertically
           ncol = 2,
           rel_widths = c(1, 1)) #number of columns in grids

```

```{r boxplots-eco-export, results = "hide"}

#Boxplot figure

fig.a <- boxplot_habs.t50 + 
  xlab("")+ 
  ylab("t50 (nm)") +
  scale_x_discrete(labels=c("scansorial" = "",
                            "other" = "")) +
  theme(legend.position = "none")

fig.b <- boxplot_act.t50 + 
  xlab("")+ 
  ylab("") +
  scale_x_discrete(labels=c("diurnal" = "",
                            "nondiurnal" = "")) +
  theme(legend.position = "none")

fig.c <- boxplot_habs.uva + 
  xlab("")+ 
  ylab("UVA (%)") +
  theme(legend.position = "none")

fig.d <- boxplot_act.uva + 
  xlab("")+ 
  ylab("") +
  theme(legend.position = "none")


# boxplot of T50 across habs
pdf("../Figures/boxplots.pdf", width = 5, height = 6)
plot_grid(fig.a, fig.b, fig.c, fig.d,
          align = 'vh', 
          labels = c("A", "B", "C", "D"),
          hjust = -0.5, 
          vjust = 1.8, 
          nrow = 2)
dev.off()

```

## Plots of t50 and %UVA values onto phylogeny

Here and in subsequent analyses, we must have one transmission value assigned to each species. We include data from both our study and from previously published literature. However, for the 8 species with data from both sources, we use our data in these visualizations and in analyses.

```{r subset-adultphy, results = "hide"}

#For species with more than one data point (measured by us and in lit), choose our data
adults_merged <- adults_sep %>%
  mutate(t50_merge = ifelse(is.na(adults_sep$t50_new), t50_lit, t50_new)) %>%
  mutate(pUVA_merge = ifelse(is.na(adults_sep$pUVA_new), pUVA_lit, pUVA_new))
```

```{r prep-ggtree, results = "hide"}

#subset data
adults_plot <- adults_merged %>%
  mutate(tip = ASW_names) %>%
  #put genus and species in separate columns
  separate(ASW_names, c("genus", "species"), sep = "_", extra = "drop") %>%
  #add tip labels with regular text
  mutate(labels = as.factor(paste(genus, species, sep = " "))) %>%
  select(tip, labels, t50_merge, pUVA_merge, hab, act)

# set row names in dataset to match the tip labels in the tree
row.names(adults_plot) <- adults_plot$tip

#check that phylogeny and data match exactly
name.check(tree.adults, adults_plot)

#resort trait dataset to the order of tree tip labels
adults_plot <- adults_plot[tree.adults$tip.label, ] 

#labels
labs <- adults_plot %>% select(tip, labels)
```

#### plot data

First we visualize the distribution of all our t50 and %UVA data across the phylogeny. 

```{r ggtree-data, fig.width = 10, fig.height=15}

# Make the phylogeny plot
p <- ggtree(tree.adults) %<+% labs + 
     geom_tiplab(cex = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(450) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (dot plot of t50 values)
p2 <- facet_plot(p, panel="t50 (nm)", data=adults_plot, geom=geom_point, aes(x=t50_merge), color = "#BDC3BE",  shape = 19, cex = 2.5)

# Make a third plot next to the first two (bar plot of %UVA)
p3 <- facet_plot(p2, panel='%UVA', data=adults_plot, geom=geom_segment, aes(x=0, xend=pUVA_merge, y=y, yend=y), color = "#BDC3BE", size=3) +
  theme_tree2(legend.position=c(.07, .93)) #add scale bars and move legends

#control widths of panels
 gt = ggplot_gtable(ggplot_build(p3))
# gtable_show_layout(gt) # will show you the layout - very handy function
# gt # see plot layout in table format
# gt$layout$l[grep('tree', gt$layout$name)] # you want to find the column specific to panel-2
 gt$widths[7] = 0.5*gt$widths[7] # reduce column 7
 gt$widths[9] = 0.5*gt$widths[9] 

#print figure
p3
```

#### plot data + scansoriality

Then we visualize our data colored by habitat. 

```{r ggtree-scansorial, fig.width = 10, fig.height=15}

# Make the phylogeny plot
p <- ggtree(tree.adults) %<+% labs + 
     geom_tiplab(cex = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(450) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (dot plot of t50 values)
p2 <- facet_plot(p, panel="t50 (nm)", data=adults_plot, geom=geom_point, aes(x=t50_merge, color = hab), shape = 19, cex = 2.5) +
  scale_color_manual(values = col_hab,
                     name = "Habitat",
                     breaks = c("scansorial", "other"),
                     labels = c("scansorial", "non-scansorial"))

# Make a third plot next to the first two (bar plot of %UVA)
p3 <- facet_plot(p2, panel='%UVA', data=adults_plot, geom=geom_segment, aes(x=0, xend=pUVA_merge, y=y, yend=y, color = hab), size=3) +
  theme_tree2(legend.position=c(.07, .93)) #add scale bars and move legends

#control widths of panels
 gt = ggplot_gtable(ggplot_build(p3))
# gtable_show_layout(gt) # will show you the layout - very handy function
# gt # see plot layout in table format
# gt$layout$l[grep('tree', gt$layout$name)] # you want to find the column specific to panel-2
 gt$widths[7] = 0.5*gt$widths[7] # reduce column 7
 gt$widths[9] = 0.5*gt$widths[9] 

#print figure
p3
```

```{r results = "hide"}
#export as pdf
pdf("../Figures/ggtree-scans.pdf", width = 10, height = 15)
grid::grid.draw(gt) # plot with grid draw
dev.off()
```

#### plot data + diurnality

Then we visualize our data colored by activity pattern. 

```{r ggtree-diurnal, fig.width = 10, fig.height=15}

# Make the phylogeny plot
p <- ggtree(tree.adults) %<+% labs + 
     geom_tiplab(cex = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(450) + 
   coord_cartesian(clip = 'off') 

# Make a second plot next to the phylogeny (dot plot of t50 values)
p2 <- facet_plot(p, panel="t50 (nm)", data=adults_plot, geom=geom_point, aes(x=t50_merge, color = act), shape = 19, cex = 2.5) +
  scale_color_manual(values = col_act,
                     name = "Activity period",
                     breaks = c("diurnal", "nondiurnal"),
                     labels = c("diurnal", "non-diurnal"))


# Make a third plot next to the first two (bar plot of %UVA)
p3 <- facet_plot(p2, panel='%UVA', data=adults_plot, geom=geom_segment, aes(x=0, xend=pUVA_merge, y=y, yend=y, color = act), size=3) +
  theme_tree2(legend.position=c(.07, .93)) #add scale bars and move legends

#control widths of panels
 gt = ggplot_gtable(ggplot_build(p3))
# gtable_show_layout(gt) # will show you the layout - very handy function
# gt # see plot layout in table format
# gt$layout$l[grep('tree', gt$layout$name)] # you want to find the column specific to panel-2
 gt$widths[7] = 0.5*gt$widths[7] # reduce column 7
 gt$widths[9] = 0.5*gt$widths[9] 

#print figure
p3
```

```{r results = "hide"}
#export as pdf
pdf("../Figures/ggtree-act.pdf", width = 10, height = 15)
grid::grid.draw(gt) # plot with grid draw
dev.off()
```


#### make paper figure

Then we visualize our data colored by ecological states that were significant in our phylogenetic ANOVAs (below). Here, we make a new column in our dataframe indicating the presence of diurnality, scansoriality, both (n = 1), and neither. This will show the distribution of both significant ecological effects, as well as show that there is only one species that is both diurnal and scansorial, indicating that these are separate effects. 

```{r, results = "hide"}
#make new category with scansorial, diurnal, both, or neither
lens_adults.new <- lens_adults.new %>%
  mutate(eco_cat = case_when(hab=="scansorial" & act=="diurnal" ~ "both",
                             hab == "scansorial" ~ "scansorial",
                             act == "diurnal" ~ "diurnal", 
                             hab=="other" & act=="nondiurnal" ~ "neither")) %>%
  mutate(eco_binary = recode_factor(eco_cat, "scansorial" = "SD", "diurnal" = "SD", "both" = "SD", "neither" = "not"))

#subset data
adults_plot.combined <- adults_plot %>%
   mutate(eco_cat = case_when(hab=="scansorial" & act=="diurnal" ~ "both",
                             hab == "scansorial" ~ "scansorial",
                             act == "diurnal" ~ "diurnal", 
                             hab == "other" & act == "nondiurnal" ~ "neither")) %>%
  select(tip, labels, t50_merge, pUVA_merge, eco_cat)

# set row names in dataset to match the tip labels in the tree
row.names(adults_plot.combined) <- adults_plot.combined$tip

#check that phylogeny and data match exactly
name.check(tree.adults, adults_plot.combined)

#resort trait dataset to the order of tree tip labels
adults_plot.combined <- adults_plot.combined[tree.adults$tip.label, ] 

#labels
labs <- adults_plot.combined %>% select(tip, labels)

#make color pallette for new combined states
col_ecocat <- c("both" = "#C107B6",
                "diurnal" = "#FFA010",
                "neither" = "black", #boxplot fig uses #636363 but with alpha it's lighter
                "scansorial" = "#009E73")

shape_ecocat <- c("both" = 23,
                "diurnal" = 25,
                "neither" = 21, 
                "scansorial" = 24)

# create vectors of colors for habitat
col_hab <- c("scansorial" = "#009E73",
              "other"  = "#636363")

#create vector of colors for activity period
col_act <- c("nondiurnal" = "#636363",
             "diurnal" = "#FFA010")
```


```{r ggtree-fig, fig.width = 10, fig.height=15}

# Make the phylogeny plot
p <- ggtree(tree.adults) %<+% labs + 
     geom_tiplab(cex = 3, aes(label = labels), fontface = 3) + 
   xlim_tree(450) + 
   coord_cartesian(clip = 'off') 


# Make a second plot next to the phylogeny (dot plot of t50 values)
p2 <- facet_plot(p, panel="t50 (nm)", data=adults_plot.combined, geom=geom_point, aes(x=t50_merge, fill = eco_cat, col = eco_cat, shape = eco_cat), cex = 2.8, alpha = 0.9) +
  scale_fill_manual(values = col_ecocat,
                     name = "Ecology",
                     breaks = c("scansorial", "diurnal", "both", "neither"),
                     labels = c("scansorial", "diurnal", "both", "neither")) +
  scale_color_manual(values = col_ecocat,
                     name = "Ecology",
                     breaks = c("scansorial", "diurnal", "both", "neither"),
                     labels = c("scansorial", "diurnal", "both", "neither")) +
  scale_shape_manual(values = shape_ecocat,
                     name = "Ecology",
                     breaks = c("scansorial", "diurnal", "both", "neither"),
                     labels = c("scansorial", "diurnal", "both", "neither"))

# Make a third plot next to the first two (bar plot of %UVA)
p3 <- facet_plot(p2, panel='%UVA', data=adults_plot.combined, geom=geom_segment, aes(x=0, xend=pUVA_merge, y=y, yend=y, color = eco_cat), size=3, alpha = 0.9) +
  theme_tree2(legend.position=c(.07, .93)) #add scale bars and move legends

#control widths of panels
 gt = ggplot_gtable(ggplot_build(p3 +# set transparency
    theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA)
        )))
# gtable_show_layout(gt) # will show you the layout - very handy function
# gt # see plot layout in table format
# gt$layout$l[grep('tree', gt$layout$name)] # you want to find the column specific to panel-2
 gt$widths[7] = 0.5*gt$widths[7] # reduce column 7
 gt$widths[9] = 0.5*gt$widths[9] 

 #print figure
p3 
```

```{r, results = "hide"}
#export as pdf
pdf("../Figures/ggtree-fig.pdf", width = 10, height = 15)
grid::grid.draw(gt) # plot with grid draw
dev.off()
```



